Equalizer 2.0\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Tweaks
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@40 db7c848b-7755-4df2-b763-8796797afa52

{\em Equalizers} are an addition to compound trees. They modify parameters of their
respective subtree at runtime to dynamically optimize one aspect of the decomposition. Due
to their nature, they are transparent to application developers, but might have
application-accessible parameters to tune their behavior.
\FIXME{ok, this is a bit tricky to Equalizer as a name for the framework as well as for this, can't we resolve this?}

@@

The discovery is implemented in a separate library, hwsd (HardWare Service
Discovery), which uses a plugin-based approach to discover GPUs for GLX, AGL or
WGL windowing systems, as well as network interfaces on Linux, Mac OS X and
Windows. Furthermore, it detects the presence of VirtualGL to allow optimal
configuration of remote visualization clusters. The resources can be discovered
on the local workstation, and through the help of a simple daemon using the
zeroconf protocol, on a set of remote nodes within a visualization cluster. A
session identifier may be used to support multiple users on a single cluster.

@@

Tide (Tiled interactive display environment) is an improved version of
DisplayCluster~\cite{DC}, providing touch-based, multi-window user interface for
high-resolution tiled display walls. Remote applications receive input events
and send pixel streams using the Deflect client library. Equalizer includes full
support, enabling application-transparent integration with Tide. When a Tide
server is configured, all output channels of a view stream in parallel to one
window on the wall. In \cite{deflect} we have shown interactive framerates for
an 24 megapixel resolution over a WAN link. Events received are translated into
Equalizer events and injected into the normal event flow, allowing seamless
application integration.

@@

RSP combines various established algorithms~\cite{adamson2004negative,
  Gau:2002:MFC:506824.506832} for multicast in an open source implementation
capable of delivering wire speed transmission rates on high-speed LAN
interfaces. In the following we will outline the RSP protocol and implementation
and motivate the design decisions. Any defaults given below are for Linux or OS
X, the Windows UDP stack requires different default values which can be found in
the implementation.

@@

Handling a smooth packet flow is critical for performance. RSP actively flow
control to advance the stream. Each incoming connection actively acknowledges
every $n$ (17 by default) packets fully received. The incoming connections shift
this acknowledgment by their connection identifier to avoid bursts of acks. Any
missed datagram is actively nack'ed as soon as it is detected. Write connections
continuously retransmit packets from nack datagrams, and advance their
\textsf{\_writeBuffers} upon reception of acks from all readers. Furthermore,
the writer will explicitly request an ack or nack when it runs out of empty
buffers or finishes its write queue.

@@

%----------------------------------------------------------------------
\appendices
% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi
%----------------------------------------------------------------------
We would like to thank and acknowledge the following institutions and projects
for providing the 3D geometry and volume test data sets: the Digital
Michelangelo Project, Stanford 3D Scanning Repository, Cyberware Inc.,
volvis.org and the Visual Human Project.  This work was partially supported by
the Swiss National Science Foundation Grant 200021-116329/1.

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/stefan}}]{Stefan Eilemann}
Stefan is the technical manager of the Visualization Team in the Blue Brain
Project. His work involves working towards large-scale visualization for
Exascale simulations; interactive integration of simulation, analysis and
visualization as well as flexible frameworks for data sharing and dynamic
allocation of heterogenous resources. He is the CEO and founder of Eyescale and
also the lead developer of the Equalizer parallel rendering framework. He has
worked in SGI's Advanced Graphics Division as the technical lead of OpenGL
Multipipe SDK - a toolkit for scalable, parallel visualization software on
shared memory supercomputers. He received his masters degree in Computer Science
from EPFL in 2015, and his Engineering Diploma in Computer Science in 1998. He
is currently working towards a PhD in Computer Science at the Visualization and
Multimedia Lab at the University of Zurich.
\end{IEEEbiography}

>>>>

{\em Equalizers} are an addition to compound trees. They modify parameters of their
respective subtree at runtime to dynamically optimize one aspect of the decomposition. Due
to their nature, they are transparent to application developers, but might have
application-accessible parameters to tune their behavior.
\FIXME{ok, this is a bit tricky to have Equalizer as a name for the framework as well as for this component, can't we change this?}

@@

The discovery is implemented in a separate library, hwsd\FIXME{need to somehow identify this term in the text,
UPPERCASE or some special font type} (HardWare Service
Discovery), which uses a plugin-based approach to discover GPUs for GLX, AGL or
WGL windowing systems, as well as network interfaces on Linux, Mac OS X and
Windows. Furthermore, it detects the presence of VirtualGL to allow optimal
configuration of remote visualization clusters. The resources can be discovered
on the local workstation, and through the help of a simple daemon using the
zeroconf protocol, on a set of remote nodes within a visualization cluster. A
session identifier may be used to support multiple users on a single cluster.

@@

\FIXME{Tide and Deflect and other 'names' should be clearly identifiable in the text.
We have many of them and sometimes this isn't visually clear in the text.}
Tide (Tiled interactive display environment) is an improved version of
DisplayCluster~\cite{DC}, providing a touch-based, multi-window user interface for
high-resolution tiled display walls. Remote applications receive input events
and send pixel streams using the Deflect client library. Equalizer includes full
support, enabling application-transparent integration with Tide. When a Tide
server is configured, all output channels of a view stream in parallel to one
window on the wall. In \cite{deflect} we have shown interactive framerates for
a 24 megapixel resolution over a WAN link. Events received are translated into
Equalizer events and injected into the normal event flow, allowing seamless
application integration.

@@

RSP combines various established algorithms~\cite{adamson2004negative,
  Gau:2002:MFC:506824.506832} for multicast in an open source implementation
capable of delivering wire speed transmission rates on high-speed LAN
interfaces. In the following we will outline the RSP protocol and implementation
as well as motivate the design decisions. Any defaults given below are for Linux or OS
X, the Windows UDP stack requires different default values which can be found in
the implementation.

@@

Handling a smooth packet flow is critical for performance. \FIXME{RSP actively flow
control to advance the stream}. Each incoming connection actively acknowledges
every $n$ (17 by default) packets fully received. The incoming connections shift
this acknowledgment by their connection identifier to avoid bursts of acks. Any
missed datagram is actively nack'ed as soon as it is detected. Write connections
continuously retransmit packets from nack datagrams, and advance their
\textsf{\_writeBuffers} upon reception of acks from all readers. Furthermore,
the writer will explicitly request an ack or nack when it runs out of empty
buffers or finishes its write queue.

@@

%----------------------------------------------------------------------
\appendices
% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi
%----------------------------------------------------------------------
We would like to thank and acknowledge the following institutions and projects
for providing the 3D geometry and volume test data sets: the Digital
Michelangelo Project, Stanford 3D Scanning Repository, Cyberware Inc.,
volvis.org and the Visual Human Project.  This work was partially supported by
the Swiss National Science Foundation Grant 200021-116329/1.\FIXME{SNF and Hasler as well as EU grants}

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/stefan}}]{Stefan Eilemann}
Stefan is the technical manager of the Visualization Team in the Blue Brain
Project. His work involves working towards large-scale visualization for
Exascale simulations; interactive integration of simulation, analysis and
visualization as well as flexible frameworks for data sharing and dynamic
allocation of heterogenous resources. He is the CEO and founder of Eyescale and
also the lead developer of the Equalizer parallel rendering framework. He has
worked in SGI's Advanced Graphics Division as the technical lead of OpenGL
Multipipe SDK - a toolkit for scalable, parallel visualization software on
shared memory supercomputers. He received his masters degree in Computer Science
from EPFL in 2015, and his Engineering Diploma in Computer Science in 1998. He
is currently working towards a PhD in Computer Science at the Visualization and
MultiMedia Lab at the University of Zurich.
\end{IEEEbiography}
